---
title: ""
author: "Tony"
date: "July 11, 2018"
output:
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  include = TRUE,
  fig.align = "center",
  results = "asis",
  fig.width = 6,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)
options(scipen = 1, digits = 2)
```

```{r config, include = FALSE}
config <-
  list(
    export_data = TRUE,
    dir_data = "data",
    export_viz = TRUE,
    dir_viz = "figs",
    download = FALSE,
    viz_footer = "By: Tony ElHabr."
  )
```

## Introduction

I wanted to do a follow-up on 
[my series of posts](https://tonyelhabr.rbind.io/posts/analysis-texas-high-school-academics-intro/) about Texas high school [University Interscholastic League](http://www.uiltexas.org/) (UIL) academic competitions
to more closely evaluate the relationship between the school performance in those
competitions with school-wide [SAT](https://en.wikipedia.org/wiki/SAT)) and 
[ACT](https://www.act.org/) scores.
For those who may not be familiar with these tests,
these are the two most popular standardized tests
used for college admission in the United States.

In my introduction to that series, I stated the following:
<i>
School-wide ... scores on state- and national-standardized 
tests (e.g. the SAT)
certainly are the most common measure of academic strength,
but I think rankings by academic competitions may be more indicative.
</i>

Essentially, I was implying that the academic UIL scores may not correspond
well, or at all, with standardized test scores.
However, I did not attempt to prove this hypothesis, 
which is what I set out to do here.
While I'm at it, I'll show the code and provide some commentary
to explain my process.

## Data Collection

While I already have collected and cleaned the UIL data that I'll need
by virtue of my work for
[my series of posts analyzing the UIL competitions](https://tonyelhabr.rbind.io/posts/analysis-texas-high-school-academics-intro/),
I did not retrieve data for standardized test scores.
To my delight,
the [Texas Education Agency's website](https://tea.texas.gov) publishes
Texas high school SAT and ACT scores for the years 2011 through 2015.
The task of scraping from this source is a perfect
use-case for the super-handy `{xml2}` and `{rvest}` packages,
as well the well-known awesome `{stringr}` and `{purrr}` packages in the `{tidyverse}`.

```{r packages}
library("tidyverse")
library("rlang")
library("teplot")
```

```{r scrape_tea, results = "hold"}
urls_tea <-
  "https://tea.texas.gov/acctres/sat_act_index.html" %>%
  xml2::read_html() %>%
  rvest::html_nodes(xpath = "//tr //td //a") %>%
  rvest::html_attr("href") %>% 
  str_subset("^\\/acctres\\/[:alpha:]{3}_[Cc]ampus_[Dd]ata")
urls_tea
```

```{r create_path_dl_tea}
create_path_dl_tea <-
  function(url_suffix = NULL, dir = "data-raw", ext = "csv") {
    if(!dir.exists(dir)) {
       dir.create(dir)
    }
    url_suffix %>%
      str_remove_all("acctres|\\/") %>% 
      paste0(".", ext) %>% 
      file.path(dir, .)
  }
```

```{r urls_tea_dl, eval = FALSE}
# NOTE(s):
# + `urls_tea_dl` is actually the same as `url_tea` because `purrr::walk()` returns its first argument.
# + `mode = "wb"` is important! Otherwise, the downloaded files have empty lines every other line
# (due to the way that CR and LFs are handled.
urls_tea_dl <-
  urls_tea %>%
  walk(
    ~download.file(
      url = paste0("https://tea.texas.gov/", .x),
      destfile = create_path_dl_tea(url_suffix = .x),
      mode = "wb"
    )
  )
```

Next, I bind the data from all of the downloaded files together and do some cleaning.

One relatively significant choice that I make is to only include the data
for the school-wide level (via the `"All Students"` filtering criteria),
although data for different demographics within each school is provided.

Additionally, there are some details of each of the standardized
tests of which the reader shooed be made aware.
The SAT has `math`, `reading`, and `writing` sections,
each having minimum and maximum scores of 200 and 800,
meaning that the `total` can range from 600 to 2400.
The ACT has `math`, `reading`, `english`, and `science` sections, 
each having a minimum and maximum score of 1 and 36,
combined for a single `compos` score also ranging from 1 to 36.
To eliminate duplicate columns representing the same underlying "thing".
I don't distinguish the `math` and `reading` section scores for each test
in separate columns, I rename the ACT's `compos` score to `total`, following
the convention used for the SAT's cumulative score.
The other sections--`writing` for the SAT and `english` and `science` for the ACT--
are not really analogous to sections in the other test, so they are filled with `NA`s
appropriately.

```{r schools_tea, results = "hide", eval = FALSE}
schools_tea <-
  urls_tea %>%
  create_path_dl_tea() %>%
  tibble(path = .) %>%
  mutate(
    test = stringr::str_extract(path, "([Ss][Aa][Tt])|([Aa][Cc][Tt])") %>% str_to_upper(),
    year = stringr::str_extract(path, "[:digit:]+") %>% as.integer()
  ) %>%
  mutate(contents =
           purrr::map2(path, test,
                       # ~import_tea_data(.x, .y)
                       ~ {
                         .x %>%
                           readr::read_csv() %>%
                           rename_all(funs(tolower)) %>%
                           filter(group == "All Students") %>%
                           select(
                             matches(
                               "name$|math|reading|writing|total|english|science|compos"
                             )
                           )
                       })) %>%
  unnest() %>%
  select(-path) %>%
  mutate_at(vars(total), funs(ifelse(test == "ACT", compos, .))) %>% 
  select(-compos) %>% 
  select(-total, everything(), total) %>% 
  rename(school = campname,
         district = distname,
         county = cntyname,
         city = regnname) %>%
  mutate_if(is.character, funs(str_replace_all(., "=|\"", ""))) %>%
  mutate_at(vars(school, district, county, city), funs(toupper)) %>% 
  # NOTE: Some county names are truncated and end with COUN or COUNT.
  # (The max seems to be 18 characters).
  # Fortunately, ther are no county names with COUN in their names, so the following
  # regular expression is sufficient.
  mutate_at(vars(county), funs(str_remove_all(., "\\s+COUN.*$"))) %>%
  # NOTE: Remove all HS/H S at the end of school names, as well as ampersands.
  # This seems to improve join percentages with other data sets.
  mutate_at(vars(school), funs(str_remove_all(., "([H]\\s*[S]$)|(\\s+\\&)") %>% str_trim())) %>%
  # NOTE: This is (try to) to resolve duplicates in raw data.
  # group_by_at(vars(matches("test|year|school|district|county|city"))) %>% 
  # summarise_all(funs(max(., na.rm = TRUE))) %>% 
  # ungroup() %>% 
  arrange(test, year, school)
schools_tea

```

```{r schools_tea_debug, include = FALSE, eval = FALSE}
# Debugging...
schools_tea
schools_tea %>% filter(str_detect(county, "COUN"))
schools_tea %>%
  count(test, year, is.na(school), sort = TRUE)
schools_tea %>%
  count(test, year, school, city, sort = TRUE) %>% 
  count(n, sort = TRUE)
schools_tea %>%
  count(test, year, school, city, sort = TRUE) %>% 
  filter(n > 1)
```

```{r schools_tea_export, include = FALSE, eval = FALSE}
teproj::export_ext_csv(
  schools_tea,
  export = config$export_data,
  dir = config$dir_data
)
```

```{r schools_tea_import, include = FALSE}
schools_tea <-
  teproj::import_ext_csv(
    schools_tea,
    dir = config$dir_data
  )
```

```{r schools_tea_show, echo = FALSE}
schools_tea %>% 
  teproj::create_kable()
```

### EDA: Year-to-Year Correlations

First, before evaluating the primary concern at hand--the relationship between
the academic UIL scores
and the SAT/ACT scores (available in the `schools_tea` data
created above)--I first want to verify that
there is some non-trivial relationship among the scores for a given school on a given
test across years. (I would be surprised if this were not shown to be true.)

```{r schools_tea_cors_byyear, results = "hide"}
schools_tea_cors_byyear <-
  schools_tea %>%
  distinct(test, year, school, .keep_all = TRUE) %>%
  filter(!is.na(total)) %>%
  unite(test_school, test, school) %>%
  widyr::pairwise_cor(
    feature = test_school,
    item = year,
    value = total
  ) %>% 
  rename(year1 = item1, year2 = item2, cor = correlation)
schools_tea_cors_byyear %>% 
  filter(year1 <= year2)
```

```{r schools_tea_cors_byyear_show, echo = FALSE}
schools_tea_cors_byyear %>% 
  filter(year1 <= year2) %>% 
  teproj::create_kable()
```

```{r visualize_cors_byyear, include = FALSE}
visualize_cors_byyear <-
  function(data = NULL,
           x = "year1",
           y = "year2",
           fill = "cor",
           alpha = 0.7,
           n_digits = 2,
           ...) {
    x <- sym(x)
    y <- sym(y)
    fill <- sym(fill)
    data %>%
      ggplot(aes(x = !!x, y = !!y, fill = !!fill)) +
      geom_tile(alpha = alpha) +
      geom_text(aes(label = round(!!fill, n_digits)), size = 5) +
      scale_fill_gradient2(...) +
      scale_x_continuous(breaks = scales::pretty_breaks()) +
      scale_y_continuous(breaks = scales::pretty_breaks()) +
      # scale_fill_viridis_c(option = "C", direction = -1, limits = limits) +
      teplot::theme_te_tile(base_size = 12) +
      coord_equal() +
      theme(legend.position = "none",
            axis.text = element_text(size = 12)) +
      labs(x = NULL,
           y = NULL)
  }
```


```{r viz_schools_tea_cors, echo = FALSE, fig.width = 8, fig.height = 8}
viz_schools_tea_cors_byyear <-
  schools_tea_cors_byyear %>%
  filter(year2 <= year1) %>% 
  visualize_cors_byyear(
    low = "white",
    mid = "yellow",
    high = "red",
    limits = c(0, 0.9),
    midpoint = 0.45
  ) +
  labs(title = str_wrap(
    paste0(
    "Correlation of SAT/ACT Scores Among Years for Texas High Schools"
    ), 80
  ),
  caption = 
    paste0(
      str_wrap(
        paste0(
          "Source: https://tea.texas.gov/acctres/sat_act_index.html."
        ),
        100
      ),
      "\n",
      config$viz_footer
    )
  )
viz_schools_tea_cors_byyear
```

As expected, there is some strong correlations among the years for school-wide scores
on these tests.

```{r viz_schools_tea_cors_export, include = FALSE}
teproj::export_ext_png(
  viz_schools_tea_cors_byyear,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 8
)
```

Ok, now let's bring in the "cleaned" school data (`schools_uil`) that I collected
and cleaned my UIL analysis. I'll subset the data to include only the same years
found in `schools_tea`--2011 through 2015.

```{r schools_uil, results = "hide"}
schools_uil <-
  file.path("data", "schools_uil.csv") %>% 
  readr::read_csv() %>% 
  filter(year >= 2011, year <= 2015)
schools_uil
```

```{r schools_uil_show, echo = FALSE}
schools_uil %>% 
  teproj::create_kable()
```

Now let's review this data set to see if year-to-year correlations also exist with it.

Importantly, some choice about how to quantify performance needs to be made.
As I discussed in
[my long-form series of posts exploring the UIL academic data](https://tonyelhabr.rbind.io/posts/analysis-texas-high-school-academics-intro/),
the evaluation of performance is somewhat subjective. Should we use
number of times a school advanced to the next level of competition in a given year?
(Note that there are three competition levels--`District`, `Region`, and `State`.)
What about the number the number of other schools it "defeated" in head-to-head competitions?
In that separate analysis, I made the choice to use the percentile rank (`prnk`) of the school's placings
across all competition levels for a given competition type (`comp`). 
I believe this measure bests represent
a school's quality of performance (where a higher value indicates better performance).
As I stated there when explaining
my choice to use percent rank for identifying "dominant" individual",
<i>
"I choose to use percent rank--which is a always a value between 0 and 1--because
it inherently accounts for the
wide range of number of competitors across all competitions. 
(For this context, a percent rank of 1 corresponds to the highest score in a given
competition, and, conversely,
a value of 0 corresponds to the lowest score.)"
</i>

Aside from this important choice regarding performance evaluation in academic UIL competitions,
note that I treat the competition type (`comp`) in `schools_uil`
as analogous to the `test` variable 
indicating SAT or ACT score in the `schools_tea` data set.
For those who have not read through my UIL analysis, note that scores for
five different competition types
was collected--`Calculator Applications`, `Computer Science`, `Mathematics`, `Number Sense`, and `Science`.


```{r schools_uil_cors_byyear, results = "hide"}
schools_uil_cors_byyear <-
  schools_uil %>% 
  select(year, school, city, comp, prnk) %>% 
  group_by(year, school, city, comp) %>% 
  summarise(prnk_sum = sum(prnk, na.rm = TRUE)) %>%
  ungroup() %>% 
  unite(comp_school, comp, school) %>%
  widyr::pairwise_cor(
    feature = comp_school,
    item = year,
    value = prnk_sum
  ) %>% 
  rename(year1 = item1, year2 = item2, cor = correlation)

schools_uil_cors_byyear %>% 
  filter(year1 <= year2) 
```

```{r schools_uil_cors_byyear_show, echo = FALSE}
schools_uil_cors_byyear %>% 
  filter(year1 <= year2) %>% 
  teproj::create_kable()
```

```{r viz_schools_uil_cors_byyear, echo = FALSE, fig.width = 8, fig.height = 8}
viz_schools_uil_cors_byyear <-
  schools_uil_cors_byyear %>% 
  filter(year2 <= year1) %>% 
  visualize_cors_byyear(
    low = "white",
    mid = "yellow",
    high = "red",
    limits = c(0, 0.9),
    midpoint = 0.45
  ) +
  labs(title = str_wrap(
    paste0(
    "Correlation of UIL Academic Scores Among Years for Texas High Schools"
    ), 80
  ),
  caption =
    paste0(
      str_wrap(
        paste0(
          "Data: Raw scores for competitions at all levels and of all types from 2011 through 2015."

        ),
        100
      ),
      "\n",
      "Source: https://www.hpscience.net/.",
      "\n",
      config$viz_footer
    )
  )
viz_schools_uil_cors_byyear
```

```{r viz_schools_uil_cors_export, include = FALSE}
teproj::export_ext_png(
  viz_schools_uil_cors_byyear,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 8
)
```

We can see that correlations among years do exist, as we would expect.
The strength of the correlations decrease for years that are farther apart,
which is also what we might expect.

### Fuzzy Joining

An important thing to keep in mind is that
the academic UIL data was scraped from a different website--https://www.hpscience.net/ (which 
was the only site that I found to have the academic competition scores)--than
that used to collect the SAT and ACT scores here. 

In its raw form,
the UIL data is much "more unclean" than the data from the TEA website, and some/many
of the names are inconsistent. In fact, the rigor involved in cleaning the UIL data
obliged be to completely hide it from the reader in my write-ups on the topic.
Nonetheless, I was able to eliminate much of the "self-inconsistency" of the UIL data
and create a suitable data set to use as one of the basis for that long-form analysis.

However, given this context--in which I am comparing data from two
different data sources--the reality is that there are more than a few differences in the
names of schools, so joining the data from the two different sources is tricky. 
To aid in my experimentation
with "fuzzy joining", I use a function from my `{tetidy}` 
package--`join_fuzzily()`, which is essentially
a wrapper for the `stringdist_*_join()` functions provided by the `{fuzzyjoin}` package.
With this function, I was able to evaluate different values for `max_dist` and
different join columns to
make a judgement regarding the quality of joins. I primarily considered counts
of joined and unjoined rows computed with
the `summarise_join_stats()` function from my `{tetidy}` package
--to make a decision on how I should join the UIL and
SAT/ACT school data. [^dry]

[^dry]:
Check out 
[my post on the "DRY" principle and its application to `R` packages](https://tonyelhabr.rbind.io/posts/dry-principle-make-a-package/).
Creating packages for actions that you perform across projects is a real time-saver!

What follows is the results of some of my experimentation.

```{r join_fuzzily}
# join_fuzzily <-
#   function(x = NULL,
#            y = NULL,
#            mode = "inner",
#            max_dist = 0,
#            cols_join = "school",
#            copy = FALSE,
#            suffix_x = "_x",
#            suffix_y = "_y") {
# 
#     f <- sprintf("fuzzyjoin::stringdist_%s_join", mode)
# 
#     ret <-
#       teproj::do_call_with(f, list(x = x, y = y, by = cols_join, max_dist = max_dist))
# 
#     ret <-
#       ret %>%
#       rename_at(vars(ends_with(".x")), funs(gsub("\\.x", suffix_x, .))) %>%
#       rename_at(vars(ends_with(".y")), funs(gsub("\\.y", suffix_y, .)))
# 
#     if(copy) {
#       cols_join_y <- quo_name(paste0(cols_join, suffix_y))
#       ret <-
#         ret %>%
#         left_join(y %>% mutate(!!!sym(cols_join_y) := !!!sym(cols_join)), by = cols_join)
#     }
#     ret
#   }
```

```{r summ_schools_joinfuzzy, results = "hide"}
schools_uil_distinct <-
  schools_uil %>%
  distinct(school, city)
```

```{r summ_schools_joinfuzzy_1, results = "hide"}
summ_schools_joinfuzzy_1 <-
  schools_tea %>%
  tetidy::join_fuzzily(
    schools_uil_distinct,
    mode = "full",
    max_dist = 0,
    cols_join = c("school"),
    suffix_x = "_tea",
    suffix_y = "_uil"
  ) %>%
  tetidy::summarise_join_stats(school_uil, school_tea) %>% 
  select(-x, -y) %>% 
  gather(metric, value)
summ_schools_joinfuzzy_1
```

```{r summ_schools_joinfuzzy_1_show, echo = FALSE}
summ_schools_joinfuzzy_1 %>% 
  teproj::create_kable()
```

```{r summ_schools_joinfuzzy_2, results = "hide"}
# NOTE: Use a full join to get more informative output from `tetidy::summarise_join_math()`.
summ_schools_joinfuzzy_2 <-
  schools_tea %>%
  unite(school_city, school, city, remove = FALSE) %>%
  tetidy::join_fuzzily(
    schools_uil_distinct %>%
      unite(school_city, school, city, remove = FALSE),
    mode = "full",
    max_dist = 0,
    cols_join = c("school_city"),
    suffix_x = "_tea",
    suffix_y = "_uil"
  ) %>%
  tetidy::summarise_join_stats(school_city_uil, school_city_tea) %>% 
  select(-x, -y) %>% 
  gather(metric, value)
summ_schools_joinfuzzy_2
```

```{r summ_schools_joinfuzzy_2_show, echo = FALSE}
summ_schools_joinfuzzy_2 %>% 
  teproj::create_kable()
```

```{r schools_join_old, results = "hide", include = FALSE, eval = FALSE}
schools_join <-
  schools_tea %>%
  unite(school_city, school, city, remove = FALSE) %>%
  tetidy::join_fuzzily(
    schools_uil_distinct %>%
      unite(school_city, school, city, remove = FALSE),
    mode = "inner",
    max_dist = 0,
    cols_join = c("school_city"),
    suffix_x = "_tea",
    suffix_y = "_uil"
  ) %>% 
  select(-matches("school_city")) %>% 
  # NOTE: Do this to remove some rows that are recorded twice in the original data.
  # Duplicates may be due to same school-city combination for two different districts.
  group_by_at(vars(matches("test|year|school|county|city"))) %>% 
  summarise_all(funs(max(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  mutate_at(vars(matches("math|reading|writing|english|science|total")),
            funs(if_else(is.infinite(.), NA_real_, .)))
schools_join
```

After some further "offline" experimentation, 
I came to conclusion that none of these fuzzy join
attempts was sufficiently adequate for my liking.
Thus, I decided that I would simply combine the two data sets 
with an `inner_join()` on
`school` and  `city` (as well as `year`, to make sure that
school scores match up properly) with the understanding that many schools end up being excluded
due to imperfect matching. With data

### "Final" Correlation Analysis

So, at this point, having decided how to join the data,
I have set myself up to do that which I set out to do--evaluate
the relationship between the academic UIL competition scores and the national SAT/ACT scores.

In order to put the two sets of data on "equal grounds",
I only evaluate math scores.
In particular, I filter `comp` in the UIL data to just the
mathematically-based competitions--`Calculator Applications`,
`Mathematics`, and `Number Sense`--excluding `Science` and `Computer Science`. And,
for the SAT/ACT data, I select only the `math` score, which
is available fore both tests, excluding
the `total` and `reading` scores also available for each and the 
`writing`, `english`, and `science` scores available for one or the other.
(Perhaps the ACT's `science` score could be compared to the `Science` UIL scores,
but I choose not to do so here.)

```{r schools_uil_math, results = "hide"}
schools_uil_math <-
  schools_uil %>%
  filter(str_detect(comp, "Calculator|Math|Number")) %>%
  group_by(year, school, city) %>% 
  summarise(prnk_sum = sum(prnk, na.rm = TRUE)) %>%
  ungroup() %>% 
  # NOTE: "Renormalize" `prnk_sum`.
  mutate(math_prnk = percent_rank(prnk_sum)) %>% 
  select(-prnk_sum)
schools_uil_math
```

```{r schools_uil_math_show, echo = FALSE}
schools_uil_math %>% 
  teproj::create_kable()
```

```{r schools_tea_math, results = "hide"}
schools_tea_math <-
  schools_tea %>%
  select(test, year, school, city, math) %>%
  filter(!is.na(math)) %>% 
  group_by(test) %>% 
  mutate(math_prnk = percent_rank(math)) %>%
  ungroup() %>% 
  group_by(year, school, city) %>% 
  summarise_at(vars(math_prnk), funs(mean(., na.rm = TRUE))) %>% 
  ungroup()
schools_tea_math
```

```{r schools_tea_math_show, echo = FALSE}
schools_tea_math %>% 
  teproj::create_kable()
```

```{r schools_join_math, results = "hide"}
schools_join_math <-
  schools_tea_math %>%
  rename_at(vars(matches("^math")), funs(paste0("tea_", .))) %>% 
  inner_join(schools_uil_math %>% 
             rename_at(vars(matches("^math")), funs(paste0("uil_", .))),
           by = c("year", "school", "city")) %>%
  select(year, school, city, matches("math"))
schools_join_math
```

```{r schools_join_math_show, echo = FALSE}
schools_join_math %>% 
  teproj::create_kable()
```

```{r schools_join_math_cors, results = "hide"}
schools_join_math_cors <-
  schools_join_math %>%
  select(-year) %>% 
  select_if(is.numeric) %>%
  corrr::correlate()
schools_join_math_cors
```

```{r schools_join_math_cors_show, echo = FALSE}
schools_join_math_cors %>% 
  teproj::create_kable()
```

```{r cor_join, include = FALSE, echo = FALSE}
cor_join <-
  schools_join_math_cors %>% 
  gather(cor, value, -rowname) %>% 
  filter(!is.na(value)) %>% 
  distinct(value) %>% 
  pull(value)
```

So, this correlation value--`r cor_join`--seems fairly low. At face value, it certainly does
not provide any basis to claim that schools that do well in the academic UIL
competitions also do well with SAT/ACT tests.
However, perhaps if I used  a different methodology, the result would be different.
Other metrics used to quantify academic UIL
performance could be tested in some kind of sensitivity analysis.

### EDA: Year-to-Year Correlations, Cont.

While I won't do any kind of rigorous second evaluation here,
I do want to try to quantify
the impact of the "missing" data dropped due to mismatched school names.
If all possible data had been used,
would the final correlation value increased (or decreased)?
A direct answer to this question is impossible, but we can evaluate the difference
in the year-to-year
correlations of scores from the schools that are joined with the
correlations calculated for 
all in "unjoined" `schools_tea` and `schools_uil` data sets. If we find
that there are large discrepancies (one way or the other), then we may have some
reason to believe that the `r cor_join` number found above is misleading.

```{r schools_postjoin_math_tidy}
schools_postjoin_math_tidy <-
  schools_join_math %>%
  unite(school_city, school, city) %>% 
  gather(metric, value, matches("prnk"))
```

```{r pairwise_cor_f}
pairwise_cor_f1 <-
  function(data = NULL, which = c("tea", "uil")) {
    which <- match.arg(which)
    data %>%
      filter(metric %>% str_detect(which)) %>% 
      # filter_at(vars(value), all_vars(!is.nan(.))) %>% 
      widyr::pairwise_cor(
        feature = school_city,
        item = year,
        value = value
      ) %>% 
      rename(year1 = item1, year2 = item2, cor = correlation) %>% 
      mutate(source = which %>% toupper())
  }

pairwise_cor_f2 <-
  function(data = NULL, which = c("tea", "uil")) {
    which <- match.arg(which)
    col <-
      data %>%
      names() %>% 
      str_subset("math")
    data %>%
      unite(school_city, school, city) %>% 
      rename(value = !!rlang::sym(col)) %>%
      mutate(source = which %>% toupper()) %>% 
      widyr::pairwise_cor(
        feature = school_city,
        item = year,
        value = value
      ) %>% 
      rename(year1 = item1, year2 = item2, cor = correlation) %>% 
      mutate(source = which %>% toupper())
  }
```

```{r schools_math_cors_byyear_diffs, results = "hide"}
schools_postjoin_math_cors_byyear <-
  bind_rows(
    schools_postjoin_math_tidy %>% 
      pairwise_cor_f1("tea"),
    schools_postjoin_math_tidy %>% 
      pairwise_cor_f1("uil")
  )

schools_prejoin_math_cors_byyear <-
  bind_rows(
    schools_tea_math %>%
      pairwise_cor_f2("tea"),
    schools_uil_math %>%
      pairwise_cor_f2("uil")
  )

schools_math_cors_byyear_diffs <-
  schools_postjoin_math_cors_byyear %>% 
  inner_join(schools_prejoin_math_cors_byyear, 
             by = c("year1", "year2", "source"), 
             suffix = c("_join", "_unjoin")) %>% 
  mutate(cor_diff = cor_join - cor_unjoin)
```

```{r schools_math_cors_byyear_diffs_wide, results = "hide"}
schools_math_cors_byyear_diffs_wide <-
  schools_math_cors_byyear_diffs %>% 
  filter(year1 <= year2) %>% 
  select(-matches("join$")) %>% 
  unite(year_pair, year1, year2) %>% 
  spread(source, cor_diff)
schools_math_cors_byyear_diffs_wide
```

```{r schools_math_cors_byyear_diffs_wide_show, echo = FALSE}
schools_math_cors_byyear_diffs_wide %>% 
  teproj::create_kable()
```


```{r viz_schools_math_cors_byyear_diffs, echo = FALSE, fig.width = 8, fig.height = 8}
viz_schools_math_cors_byyear_diffs <-
  schools_math_cors_byyear_diffs %>%
  filter(year2 <= year1) %>% 
  visualize_cors_byyear(
    fill = "cor_diff",
    low = "cyan",
    high = "red",
    mid = "white",
    limits = c(-0.3, 0.3),
    midpoint = 0
  ) +
  facet_wrap( ~ source) +
  labs(title = str_wrap(
    paste0(
      "Difference in Year-to-Year Correlation of UIL Academic Math Scores and SAT/ACT Math Scores",
      "for Pre- vs. Post-Joined Data")
    , 80),
    caption =
      paste0(
      str_wrap(
        paste0(
          "Positive values indicate that the correlation is more positive in the post-joined data ",
          "(which has less records due to mismatched rows that are dropped).",
          "Scores for each source are rescaled to a percentile rank, where 0 is the minimum ",
          "and 1 is the maximum."
        ),
        100
      ),
      "\n",
      config$viz_footer
      )
  )
viz_schools_math_cors_byyear_diffs
```

```{r viz_schools_math_cors_byyear_diffs_export, include = FALSE}
teproj::export_ext_png(
  viz_schools_math_cors_byyear_diffs,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 8
)
```

Note that the correlations in the joined data are a bit "stronger"--in the sense that they 
are more positive--among the TEA
SAT/ACT-related data, although not in any kind of magnificent way. 
Additionally, the differences for the UIL data are trivial.
Thus, we might say that the additional data that could have possibly increased (or decreased)
the correlation value found--`r cor_join`--would not have changed much at all.

```{r schools_tea_eda, include = FALSE, eval = FALSE}
# ## Bonus Data Exploration
# 
# As a final "aside" to the focus of this post, I thought it would be interesting
# (maybe only to me) to compare the SAT and ACT scores of my 
# high school--`SAMUEL CLEMENS` in the TEA data--with our primary rival--`BYRON P STEELE II`.
# (After all, I have made the effort to extract and clean this data, so I would
# be foolish not to explore any questions that I have in mind.)
# schools_tea %>%
#   select(test, year, school, total) %>%
#   distinct(test, year, school, .keep_all = TRUE) %>%
#   filter(str_detect(school, "CLEMENS|STEELE II")) %>%
#   spread(year, total)
# Even though their football team may have been better than ours while I was attending
# high school--which is a big deal in Texas--I can say that our school 
# was better (marginally) according to a simple comparison of scores on these tests.
# That may not be saying a lot to someone with no ties to this setting, but
# it sparks a bit of pride in me!
```


## Conclusion

So, my initial inclination
[in my analysis of academic UIL competitions](https://tonyelhabr.rbind.io/posts/analysis-texas-high-school-academics-intro/)) seems correct--there is no significant relationship between
Texas high school academic competition scores and standardized test scores (for math, between 2011 and 2015).
And, with that question answered,
I intend to explore this rich data set in other ways in future blog posts.

